---
output:
  pdf_document: default
  html_document: default
title: "Bovine TB: Exploratory Data Analysis"
author: "Benjamen Simon"
date: "February 2023"
knit: (function(inputFile, encoding) { 
          rmarkdown::render(
            inputFile,
            encoding = encoding, 
            output_dir = "../Reports") 
            })
header-includes:
  - \usepackage{amsmath}
  - \usepackage{enumerate}
  - \usepackage{bm}
  - \usepackage[T1]{fontenc}
---

# Preamble

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, message = FALSE)

options(scipen=999)
```

This file is used to process the Bovine Tuberculosis data from the APHA, aggregating it to the farm and week level, for use in the inference model. This file focuses on Cheshire.

The packages we will use:

```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)
library(RPostgreSQL)
library(lubridate)
library(sf)
```

Connecting to the database

```{r Connecting to the DB, echo= FALSE, eval= FALSE}

# Save the password that we can "hide" it as best as we can by collapsing it
pw <- {
  "BtbPhD2019"
}
 
# Loads the PostgreSQL driver
drv <- dbDriver("PostgreSQL")

# Creates a connection to the postgres database
# Note that "con" will be used later in each connection to the database
con <- dbConnect(drv, dbname = "uk_btb",
                 host = "fhm-chicas-gis.lancs.ac.uk", port = 5432,
                 user = "simonb", password = pw)

rm(pw) # Removes the password
 
# Check for the locations_tbl
dbExistsTable(con, "locations_tbl")
# TRUE

```


# Pull Data

## Historic Herd Size

```{r Cheshire Data: Historical Herd, eval = FALSE}

df_historical_herd = dbGetQuery(con, 
                                    "SELECT hist.cph, loc.county, loc.parish, loc.holding, loc.county_name, hist.year, hist.month, hist.num_animals
                                     FROM historic_herd_tbl hist
                                     INNER JOIN (SELECT DISTINCT cph, county, parish, holding, county_name
                                                    FROM locations_extended_tbl
                                                 ) loc
                                             ON (hist.cph = loc.cph)
                                    ")

```

```{r}

df_historical_herd_clean <-
  df_historical_herd %>% 
  group_by(cph, county, parish, holding, county_name, year, month) %>% 
  summarise(num_animals = sum(num_animals)) %>% 
  ungroup()

```

```{r}
rm(df_historical_herd)
```


## Movements

```{r}

df_unique_movements_per_animal = dbGetQuery(con, 
                                            "SELECT eartag, COUNT(*) as num_movements
                                            FROM (
                                                   SELECT eartag, movement_date, movement_id, 
                                                          off_cph, off_county, off_parish, off_holding, off_county_name, 
                                                          on_cph, on_county, on_parish, on_holding, on_county_name, 
                                                          movement_date_day, movement_date_month, movement_date_year, movement_date_dow
                                                   FROM moves_no_dups_with_sameday_death_tbl
                                            ) df_unique_movements
                                            GROUP BY eartag
                                            ")
```

```{r}

df_unique_movements_on_off_per_county = dbGetQuery(con, 
                                                    "SELECT t1.off_county_name AS county_name, t1.count AS off_count, t2.count as on_count
                                                    FROM 
                                                      (SELECT off_county_name, COUNT(*) as count FROM moves_no_dups_with_sameday_death_tbl GROUP BY off_county_name) as t1
                                                      JOIN
                                                      (SELECT on_county_name, COUNT(*) as count FROM moves_no_dups_with_sameday_death_tbl GROUP BY on_county_name) as t2
                                                      ON t1.off_county_name = t2.on_county_name
                                                    ORDER BY county_name
                                                    ")
```



## Births and Deaths

```{r}

df_unique_births = dbGetQuery(con, 
    "SELECT on_county_name AS county_name, movement_date_year AS year, movement_date_month AS month, COUNT(*) as num_births
                                            FROM (
                                                   SELECT eartag, movement_date, 
                                                          off_cph, off_county, off_parish, off_holding, off_county_name, 
                                                          on_cph, on_county, on_parish, on_holding, on_county_name, 
                                                          movement_date_day, movement_date_month, movement_date_year, movement_date_dow
                                                   FROM births_no_dups_tbl
                                            ) df_unique_births
    GROUP BY movement_date_month, movement_date_year, on_county_name
    ORDER BY on_county_name, movement_date_year, movement_date_month
     "
    )

```


```{r}

df_unique_deaths = dbGetQuery(con, 
    "SELECT off_county_name AS county_name, movement_date_year AS year, movement_date_month AS month, COUNT(*) as num_deaths
                                            FROM (
                                                   SELECT eartag, movement_date, 
                                                          off_cph, off_county, off_parish, off_holding, off_county_name, 
                                                          on_cph, on_county, on_parish, on_holding, on_county_name, 
                                                          movement_date_day, movement_date_month, movement_date_year, movement_date_dow
                                                   FROM deaths_no_dups_tbl
                                            ) df_unique_deaths
    GROUP BY movement_date_month, movement_date_year, off_county_name
    ORDER BY off_county_name, movement_date_year, movement_date_month
     "
    )

```


## Movement patterns in time

```{r}

df_unique_movements_per_dow = dbGetQuery(con, 
                                            "SELECT movement_date_dow, movement_date_month, movement_date_year, COUNT(*) as num_movements
                                            FROM (
                                                   SELECT eartag, movement_date, movement_id, 
                                                          off_cph, off_county, off_parish, off_holding, off_county_name, 
                                                          on_cph, on_county, on_parish, on_holding, on_county_name, 
                                                          movement_date_day, movement_date_month, movement_date_year, movement_date_dow
                                                   FROM moves_no_dups_with_sameday_death_tbl
                                            ) df_unique_movements
                                            GROUP BY movement_date_dow, movement_date_month, movement_date_year
                                            ")
```



## Testing

```{r}

df_total_tests_dates = dbGetQuery(con, 
    "SELECT test_date_dow, test_date_month, test_date_year, COUNT(*)
     FROM tests_no_dups_extended
     GROUP BY test_date_dow, test_date_month, test_date_year
    ")

```


```{r}

df_total_tests_by_category = dbGetQuery(con, 
    "SELECT category, test_date_year, COUNT(*)
     FROM tests_no_dups_extended
     GROUP BY category, test_date_year
    ")

```

```{r}

df_total_tests_by_action = dbGetQuery(con, 
    "SELECT category, action, test_date_year, COUNT(*)
     FROM tests_no_dups_extended
     GROUP BY category, action, test_date_year
    ")

```



Load the shape files:

```{r}

# Parish boundaries
APHA_gdb <- sf::st_read(dsn = "Spatial/APHA_England_Wales_Parishes.gdb/APHA_Parishes.gdb")

APHA_shp <- sf::st_read(dsn = "Spatial/APHA_England_Wales_Parishes_shp/APHA_England_Wales_Parishes.shp")

```

```{r}

county_shape_data <- APHA_gdb %>%
  group_by(COUNTY, C_NUM) %>%
  summarize() %>%
  st_as_sf()

```


```{r}
# Rows to combine (names "London" and "East London")
names_to_combine <- c("GREATER LONDON", "GREATER LONDON - EAST")

# Filter the rows with the specified names
rows_to_combine <- county_shape_data %>% filter(COUNTY %in% names_to_combine)

GL_row = county_shape_data[(county_shape_data$COUNTY == "GREATER LONDON"), ]

GL_row$Shape = st_union(rows_to_combine$Shape)

# Filter out the original rows that were combined
county_shape_data_modified <- county_shape_data %>% filter(!COUNTY %in% names_to_combine)

# Add the new row to the original data frame
county_shape_data_modified <- rbind(county_shape_data_modified, GL_row)

```


```{r}
# Rows to combine (names "London" and "East London")
names_to_combine <- c("HEREFORDSHIRE, COUNTY OF", "WORCESTERSHIRE")

# Filter the rows with the specified names
rows_to_combine <- county_shape_data %>% filter(COUNTY %in% names_to_combine)

HW_row = county_shape_data[(county_shape_data$COUNTY == "HEREFORDSHIRE, COUNTY OF"), ]

HW_row$COUNTY = "HEREFORD & WORCESTER"
HW_row$Shape = st_union(rows_to_combine$Shape)

# Filter out the original rows that were combined
county_shape_data_modified <- county_shape_data_modified %>% filter(!COUNTY %in% names_to_combine)

# Add the new row to the original data frame
county_shape_data_modified <- rbind(county_shape_data_modified, HW_row)

```


```{r}
# Rows to combine (names "London" and "East London")
names_to_combine <- c("NORTH POWYS", "SOUTH POWYS")

# Filter the rows with the specified names
rows_to_combine <- county_shape_data %>% filter(COUNTY %in% names_to_combine)

PW_row = county_shape_data[(county_shape_data$COUNTY == "NORTH POWYS"), ]

PW_row$COUNTY = "POWYS"
PW_row$Shape = st_union(rows_to_combine$Shape)

# Filter out the original rows that were combined
county_shape_data_modified <- county_shape_data_modified %>% filter(!COUNTY %in% names_to_combine)

# Add the new row to the original data frame
county_shape_data_modified <- rbind(county_shape_data_modified, PW_row)

```

```{r}

county_shape_data_modified[(county_shape_data_modified$C_NUM == 50), ]$COUNTY = "NORTH YORKSHIRE (SOUTHERN)"

```



```{r}
county_shape_data_modified <- county_shape_data_modified %>%
  mutate(COUNTY = case_when(
    COUNTY == "DEVONSHIRE" ~ "DEVON",
    COUNTY == "GREATER MANCHESTER" ~ "GREATER  MANCHESTER",
    COUNTY == "HERTFORD" ~ "HERTFORDSHIRE",
    COUNTY == "LEICESTER" ~ "LEICESTERSHIRE",
    TRUE ~ COUNTY 
  ))


county_shape_data_modified <-  county_shape_data_modified[order(county_shape_data_modified$COUNTY), ]
```

```{r}
county_shape_data_modified <- county_shape_data_modified %>% st_as_sf()
```

```{r}
rm(APHA_gdb, APHA_shp)
rm(rows_to_combine, GL_row, HW_row, PW_row, names_to_combine)
rm(county_shape_data)
```



```{r}
#map_names_plot <- 
  ggplot(data = county_shape_data_modified) +
  geom_sf(aes(fill = C_NUM)) +
  scale_fill_viridis_c(option = "plasma", trans = "sqrt") + 
  geom_sf_text(aes(label = COUNTY), fun.geometry = sf::st_centroid) + 
  theme_minimal()
```


```{r}

ggsave("map_names_plot.png", plot = map_names_plot, width = 20, height = 40, units = "in", dpi = 300)

```







# Historic Herd

## Total number of animals per month per year


```{r}

HH_plot <-
  df_historical_herd_clean %>% 
  mutate_at(vars(-num_animals), as.factor) %>% 
  group_by(year, month) %>%
  summarize(counts = sum(num_animals)) %>% 
  ggplot(aes(x = month, y = counts, group = year, color = year)) +
  geom_line()+
  labs(x = "Month", y = "Count", color = "Year")+
  theme_minimal() +
  labs(title = "Total number of animals in England and Wales", x = "Month", y = "Total number of animals")

```

```{r}

ggsave("Figures/Total number of animals in England and Wales.png", plot = HH_plot, width = 12, height = 6, units = "in", dpi = 300)

```


## Number of animals per farm (banded)

```{r}
df_hist_herd_grouped <-
  df_historical_herd_clean %>% 
  mutate(bands = cut(num_animals, breaks = c(0, 5, 10, 20, 50, 100, 200, 500, 750, 1000, 2000, 5000, Inf), labels = c("1-5", "6-10", "11-20", "21-50", "51-100", "101-200", "201-500", "501-750", "751-1000", "1001-2000", "2001-5000", "5000+"))) %>% 
  mutate(bands = as.factor(bands))

```


```{r}
HH_grouped_plot <-
  ggplot(df_hist_herd_grouped, aes(x = bands)) +
  geom_bar(color="darkblue", fill = "lightblue")+
  theme_minimal() +
  theme(axis.text.x = element_text(angle = -45, hjust = 0)) +
  labs(title = "Average number of animals per farm", x = "Number of animals", y = "Number of observations (farms by month)")
```

```{r}

ggsave("Figures/Average number of animals per farm.png", plot = HH_grouped_plot, width = 12, height = 6, units = "in", dpi = 300)

```


## Average number of animals per county

```{r}
df_historical_herd_summary <-
  df_historical_herd_clean %>%
  group_by(county_name, month, year) %>%
  summarize(total_animals = sum(num_animals)) %>%
  group_by(county_name) %>%
  summarize(avg_animals = mean(total_animals), se = sd(total_animals) / sqrt(n()))
```

```{r}

HH_by_county_plot <-
  ggplot(df_historical_herd_summary, aes(x = county_name, y = avg_animals)) +
  geom_bar(stat = "identity", color="darkblue", fill = "lightblue") +
  geom_errorbar(aes(ymin = avg_animals - se, ymax = avg_animals + se), width = 0.2) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  xlab("County") +
  ylab("Average Number of Animals")+
  labs(title = "Average number of animals per county")
```


```{r}

ggsave("Figures/Average number of animals per county.png", plot = HH_by_county_plot, width = 12, height = 6, units = "in", dpi = 300)

```



```{r}

Avg_animals_county_spatial_plot_labels <-
  county_shape_data %>% 
  left_join(df_historical_herd_summary, by = c('COUNTY' = 'county_name')) %>% 
  ggplot() +
  geom_sf(aes(fill = avg_animals)) +
  scale_fill_viridis_c(option = "plasma", name = "Avg. Number of Animals") + 
  geom_sf_label(aes(label = COUNTY)) + 
  theme_minimal() +
  labs(title = "Avg. Number of Animals per County")

```

```{r}

# Avg_animals_county_spatial_plot <-
  county_shape_data_modified %>% 
  left_join(df_historical_herd_summary, by = c('COUNTY' = 'county_name')) %>% 
  ggplot() +
  geom_sf(aes(fill = avg_animals)) +
  scale_fill_viridis_c(option = "plasma", name = "Avg. Number of Animals") + 
  theme_minimal() +
  labs(title = "Avg. Number of Animals per County")

```

```{r}

ggsave("Figures/Average number of animals Spatial Labels.png", plot = Avg_animals_county_spatial_plot_labels, width = 18, height = 18, units = "in", dpi = 300)

```

```{r}

ggsave("Figures/Average number of animals Spatial.png", plot = Avg_animals_county_spatial_plot, width = 6, height = 6, units = "in", dpi = 300)

```


# Movements

## Number of movements per animal (grouped)

```{r}

df_unique_movements_banded <-
  df_unique_movements_per_animal %>% 
  mutate(bands = cut(num_movements, breaks = c(0, 1, 2, 3, 4, 5, 10, 20, 50, 100, Inf), labels = c("1", "2", "3", "4", "5", "6-10", "11-20", "21-50", "51-100", "100+"))) %>% 
  mutate(bands = as.factor(bands))

```

```{r}

Movements_plot <-
  ggplot(df_unique_movements_banded, aes(x = bands)) +
  geom_bar(color="darkblue", fill = "lightblue")+
  theme_minimal() +
  theme(axis.text.x = element_text(angle = -45, hjust = 0)) +
  labs(title = "Total number of movements per animal",x = "Number of movements", y = "Number of animals")

```

```{r}

ggsave("Figures/Total number of movements per animal.png", plot = Movements_plot, width = 12, height = 6, units = "in", dpi = 300)

```


## Movements out of each county

```{r}

# convert data to long format
df_unique_movements_on_off_per_county_long <- tidyr::gather(df_unique_movements_on_off_per_county, key = "count_type", value = "count", -county_name)

```

```{r}

Movements_by_county_plot <-
  ggplot(df_unique_movements_on_off_per_county_long, aes(x = county_name, y = count, fill = count_type)) +
  geom_bar(position = "dodge", stat = "identity") +
  scale_fill_manual(values = c("#F8766D", "#00BFC4"), labels = c("Off", "On"))+
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Total number of movement on and off per county", x = "County", y = "Number of movements")

```

```{r}

ggsave("Figures/Total number of movement on and off per county.png", plot = Movements_by_county_plot, width = 12, height = 6, units = "in", dpi = 300)

```


```{r}

Moves_off_county_spatial_plot_labels <-
  county_shape_data %>% 
  left_join(df_unique_movements_on_off_per_county, by = c('COUNTY' = 'county_name')) %>% 
  ggplot() +
  geom_sf(aes(fill = off_count)) +
  scale_fill_viridis_c(option = "plasma", name = "Total Movements Off") + 
  geom_sf_label(aes(label = COUNTY)) + 
  theme_minimal() +
  labs(title = "Total Movements Off per County")

```

```{r}

#Moves_off_county_spatial_plot <-
  county_shape_data_modified %>% 
  left_join(df_unique_movements_on_off_per_county, by = c('COUNTY' = 'county_name')) %>% 
  ggplot() +
  geom_sf(aes(fill = off_count)) +
  scale_fill_viridis_c(option = "plasma", name = "Total Movements Off") + 
  theme_minimal() +
  labs(title = "Total Movements Off per County")

```

```{r}

ggsave("Figures/Total Moves Off per County Spatial Labels.png", plot = Moves_off_county_spatial_plot_labels, width = 18, height = 18, units = "in", dpi = 300)

```

```{r}

ggsave("Figures/Total Moves Off per County Spatial.png", plot = Moves_off_county_spatial_plot, width = 6, height = 6, units = "in", dpi = 300)

```



```{r}

Moves_on_county_spatial_plot_labels <-
  county_shape_data %>% 
  left_join(df_unique_movements_on_off_per_county, by = c('COUNTY' = 'county_name')) %>% 
  ggplot() +
  geom_sf(aes(fill = on_count)) +
  scale_fill_viridis_c(option = "plasma", name = "Total Movements On") + 
  geom_sf_label(aes(label = COUNTY)) + 
  theme_minimal() +
  labs(title = "Total Movements On per County")

```

```{r}

# Moves_on_county_spatial_plot <-
  county_shape_data_modified %>% 
  left_join(df_unique_movements_on_off_per_county, by = c('COUNTY' = 'county_name')) %>% 
  ggplot() +
  geom_sf(aes(fill = on_count)) +
  scale_fill_viridis_c(option = "plasma", name = "Total Movements On") + 
  theme_minimal() +
  labs(title = "Total Movements On per County")

```

```{r}

ggsave("Figures/Total Moves On per County Spatial Labels.png", plot = Moves_on_county_spatial_plot_labels, width = 18, height = 18, units = "in", dpi = 300)

```

```{r}

ggsave("Figures/Total Moves On per County Spatial.png", plot = Moves_on_county_spatial_plot, width = 6, height = 6, units = "in", dpi = 300)

```


## Births and Deaths


```{r}


```


```{r}

Births_plot <-
  df_unique_births %>% 
  group_by(month, year) %>% 
  summarise(total_births = sum(num_births)) %>% 
  arrange(year, month) %>% 
  ggplot(aes(x = factor(month), y = total_births, group = factor(year), color = factor(year))) +
  geom_line()+
  labs(x = "Month", y = "Count", color = "Year")+
  theme_minimal() +
  labs(title = "Total Births per Month", x = "Month", y = "Total number of Births")

```

```{r}

ggsave("Figures/Total Births per Month.png", plot = Births_plot, width = 12, height = 6, units = "in", dpi = 300)

```


```{r}

Deaths_plot <-
  df_unique_deaths %>% 
  group_by(month, year) %>% 
  summarise(total_deaths = sum(num_deaths)) %>% 
  arrange(year, month) %>% 
  ggplot(aes(x = factor(month), y = total_deaths, group = factor(year), color = factor(year))) +
  geom_line()+
  labs(x = "Month", y = "Count", color = "Year")+
  theme_minimal() +
  labs(title = "Total Deaths per Month", x = "Month", y = "Total number of Deaths")

```

```{r}

ggsave("Figures/Total Deaths per Month.png", plot = Deaths_plot, width = 12, height = 6, units = "in", dpi = 300)

```



```{r}

 Births_county_plot <-
  df_unique_births %>% 
  na.omit() %>% 
  group_by(county_name) %>% 
  summarise(total_births = sum(num_births)) %>% 
  ggplot(aes(x = county_name, y = total_births)) +
  geom_bar(stat = "identity", color="darkblue", fill = "lightblue") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  xlab("County") +
  ylab("Total Births") + 
  labs(title = "Total Births per County")

```

```{r}

ggsave("Figures/Total Births per County.png", plot = Births_county_plot, width = 12, height = 6, units = "in", dpi = 300)

```

```{r}
df_unique_births_per_county <-
  df_unique_births %>% 
  na.omit() %>% 
  group_by(county_name) %>% 
  summarise(total_births = sum(num_births))
```

```{r}

Births_county_spatial_plot_labels <-
  county_shape_data_modified %>% 
  left_join(df_unique_births_per_county, by = c('COUNTY' = 'county_name')) %>% 
  ggplot() +
  geom_sf(aes(fill = total_births)) +
  scale_fill_viridis_c(option = "plasma", name = "Total Births") + 
  geom_sf_label(aes(label = COUNTY)) + 
  theme_minimal() +
  labs(title = "Total Births per County")

```

```{r}

# Births_county_spatial_plot <-
  county_shape_data_modified %>% 
  left_join(df_unique_births_per_county, by = c('COUNTY' = 'county_name')) %>% 
  ggplot() +
  geom_sf(aes(fill = total_births)) +
  scale_fill_viridis_c(option = "plasma", name = "Total Births") + 
  theme_minimal() +
  labs(title = "Total Births per County")

```

```{r}

ggsave("Figures/Total Births per County Spatial Labels.png", plot = Births_county_spatial_plot_labels, width = 18, height = 18, units = "in", dpi = 300)

```

```{r}

ggsave("Figures/Total Births per County Spatial.png", plot = Births_county_spatial_plot, width = 6, height = 6, units = "in", dpi = 300)

```



```{r}

Deaths_county_plot <-
  df_unique_deaths %>% 
  na.omit() %>% 
  group_by(county_name) %>% 
  summarise(total_deaths = sum(num_deaths)) %>% 
  ggplot(aes(x = county_name, y = total_deaths)) +
  geom_bar(stat = "identity", color="darkblue", fill = "lightblue") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  xlab("County") +
  ylab("Total Deaths") + 
  labs(title = "Total Deaths per County")

```

```{r}

ggsave("Figures/Total Deaths per County.png", plot = Deaths_county_plot, width = 12, height = 6, units = "in", dpi = 300)

```

```{r}
df_unique_deaths_per_county <-
  df_unique_deaths %>% 
  na.omit() %>% 
  group_by(county_name) %>% 
  summarise(total_deaths = sum(num_deaths))
```

```{r}

Deaths_county_spatial_plot_labels <-
  county_shape_data %>% 
  left_join(df_unique_deaths_per_county, by = c('COUNTY' = 'county_name')) %>% 
  ggplot() +
  geom_sf(aes(fill = total_deaths)) +
  scale_fill_viridis_c(option = "plasma", name = "Total Deaths") + 
  geom_sf_label(aes(label = COUNTY)) + 
  theme_minimal() +
  labs(title = "Total Deaths per County")

```

```{r}

# Deaths_county_spatial_plot <-
  county_shape_data_modified %>% 
  left_join(df_unique_deaths_per_county, by = c('COUNTY' = 'county_name')) %>% 
  ggplot() +
  geom_sf(aes(fill = total_deaths)) +
  scale_fill_viridis_c(option = "plasma", name = "Total Deaths") + 
  theme_minimal() +
  labs(title = "Total Deaths per County")

```

```{r}

ggsave("Figures/Total Deaths per County Spatial Labels.png", plot = Deaths_county_spatial_plot_labels, width = 18, height = 18, units = "in", dpi = 300)

```

```{r}

ggsave("Figures/Total Deaths per County Spatial.png", plot = Deaths_county_spatial_plot, width = 6, height = 6, units = "in", dpi = 300)

```





## Movement patterns in time

```{r}

df_unique_movements_by_month <-
  df_unique_movements_per_dow %>% 
  group_by(movement_date_month, movement_date_year) %>% 
  summarise(total_movements = sum(num_movements)) %>% 
  arrange(movement_date_year, movement_date_month)

```


```{r}

Movements_by_month <-
  df_unique_movements_by_month %>% 
  ggplot(aes(x = factor(movement_date_month), y = total_movements, group = factor(movement_date_year), color = factor(movement_date_year))) +
  geom_line()+
  labs(x = "Month", y = "Count", color = "Year")+
  theme_minimal() +
  labs(title = "Total Movements per Month", x = "Month", y = "Total number of Movements")

```


```{r}

ggsave("Figures/Total Movements per Month.png", plot = Movements_by_month, width = 12, height = 6, units = "in", dpi = 300)

```


```{r}

Movements_by_dow <-
  df_unique_movements_per_dow %>% 
  group_by(movement_date_dow) %>% 
  summarise(total_moves = sum(num_movements)) %>% 
  mutate(movement_date_dow = 
           stringr::str_remove_all(movement_date_dow, pattern = ' '), 
         movement_date_dow = as.factor(movement_date_dow)) %>% 
  mutate(movement_date_dow = fct_relevel(
    movement_date_dow, 
    c("monday", "tuesday", "wednesday", "thursday", "friday", "saturday", "sunday"))) %>% 
  ggplot(aes(x = movement_date_dow, y = total_moves)) +
  geom_bar(stat = "identity", color="darkblue", fill = "lightblue") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  xlab("Day of the Week") +
  ylab("Total Movements") +
  labs(title = "Total Movements per Day of the Week")

```

```{r}

ggsave("Figures/Total Movements per Day of the Week.png", plot = Movements_by_dow, width = 12, height = 6, units = "in", dpi = 300)

```


## Testing

```{r}

df_unique_tests_by_month <-
  df_total_tests_dates %>% 
  group_by(test_date_month, test_date_year) %>% 
  summarise(total_tests = sum(count)) %>% 
  arrange(test_date_year, test_date_month)

```


```{r}

Tests_per_month <-
  df_unique_tests_by_month %>% 
  ggplot(aes(x = factor(test_date_month), y = total_tests, group = factor(test_date_year), color = factor(test_date_year))) +
  geom_line()+
  labs(x = "Month", y = "Count", color = "Year")+
  theme_minimal() +
  labs(title = "Total Tests per Month", x = "Month", y = "Total number of Tests")

```

```{r}

ggsave("Figures/Total Tests per Month.png", plot = Tests_per_month, width = 12, height = 6, units = "in", dpi = 300)

```


```{r}

Tests_per_dow <-
  df_total_tests_dates %>% 
  group_by(test_date_dow) %>% 
  summarise(total_tests = sum(count)) %>% 
  mutate(test_date_dow = 
           stringr::str_remove_all(test_date_dow, pattern = ' '), 
         test_date_dow = as.factor(test_date_dow)) %>% 
  mutate(test_date_dow = fct_relevel(
    test_date_dow, 
    c("monday", "tuesday", "wednesday", "thursday", "friday", "saturday", "sunday"))) %>% 
  ggplot(aes(x = test_date_dow, y = total_tests)) +
  geom_bar(stat = "identity", color="darkblue", fill = "lightblue") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  xlab("Day of the Week") +
  ylab("Total Tests") + 
  labs(title = "Total Tests per Day of the Week")

```

```{r}

ggsave("Figures/Total Tests per Day of the Week.png", plot = Tests_per_dow, width = 12, height = 6, units = "in", dpi = 300)

```



```{r}

Tests_per_category_per_year <-
  df_total_tests_by_category %>% 
  na.omit() %>% 
  ggplot(aes(x = test_date_year, y = count, fill = category)) +
  geom_col(position = "dodge", width = 0.8) +
  scale_fill_discrete(name = "Category") +
  scale_x_continuous(name = "Year") +
  scale_y_continuous(name = "Number of Tests") +
  theme_minimal() + 
  labs(title = "Total Tests per Category per Year")

```

```{r}

ggsave("Figures/Total Tests per Category per Year.png", plot = Tests_per_category_per_year, width = 12, height = 6, units = "in", dpi = 300)

```


```{r}

Tests_results_per_category <-
  df_total_tests_by_action %>% 
  na.omit() %>%
  ggplot(aes(x = category, y = count, fill = action)) +
  geom_col(position = "dodge", width = 0.8) +
  scale_fill_discrete(name = "Category") +
  scale_y_continuous(name = "Number of Tests") +
  theme_minimal() +
  labs(title = "Total Test Results per Category")

```

```{r}

ggsave("Figures/Total Test Results per Category.png", plot = Tests_results_per_category, width = 12, height = 6, units = "in", dpi = 300)

```


```{r}

Tests_results_per_category_exc_N <-
  df_total_tests_by_action %>% 
  na.omit() %>%
  filter(action != "N") %>% 
  ggplot(aes(x = category, y = count, fill = action)) +
  geom_col(position = "dodge", width = 0.8) +
  scale_fill_discrete(name = "Category") +
  scale_y_continuous(name = "Number of Tests") +
  theme_minimal() + 
  labs(title = "Total Test Results per Category (exc. Negative)")

```

```{r}

ggsave("Figures/Total Test Results per Category exc N.png", plot = Tests_results_per_category_exc_N, width = 12, height = 6, units = "in", dpi = 300)

```




































































