---
output:
  pdf_document: default
  html_document: default
title: "Bovine TB: Creating SQL Tables"
author: "Benjamen Simon"
date: "November 2022"
knit: (function(inputFile, encoding) { 
          rmarkdown::render(
            inputFile,
            encoding = encoding, 
            output_dir = "../Reports") 
            })
header-includes:
  - \usepackage{amsmath}
  - \usepackage{enumerate}
  - \usepackage{bm}
  - \usepackage[T1]{fontenc}
---


# Preamble

This file is used to create and extract tables from the Bovine TB database I have set up, using SQL queries.

The packages we will use:

```{r setup, include=FALSE}
library("RPostgreSQL") # Used to connect to and query the database

library(tidyverse)
library(ggplot2)
```

Connecting to the database:

```{r Connecting to the DB, echo= FALSE}
 
# Create a connection
# Save the password that we can "hide" it as best as we can by collapsing it
pw <- {
  "BtbPhD2019"
}
 
# Loads the PostgreSQL driver
drv <- dbDriver("PostgreSQL")

# Creates a connection to the postgres database
# Note that "con" will be used later in each connection to the database
con <- dbConnect(drv, dbname = "uk_btb",
                 host = "fhm-chicas-gis.lancs.ac.uk", port = 5432,
                 user = "simonb", password = pw)

rm(pw) # Removes the password
 
# Check for the locations_tbl
dbExistsTable(con, "locations_tbl")
# TRUE

```

```{r, Disconnecting from the DB, echo= FALSE, eval = FALSE}
# Close the connection
# Run at the end
dbDisconnect(con)
dbUnloadDriver(drv)
```



# Locations

## Counties, Parishes, and Holdings

### The number of parishes and holdings in each county:

```{r Query: Counties structure, eval = FALSE}

df_count_parishes_and_holdings_per_county = dbGetQuery(con, 
    "SELECT loc2.county_name, loc2.county, count(loc2.parish) AS num_parishes, SUM(loc2.num_holdings_in_parish) AS num_holdings
     FROM (
    		    SELECT loc.county_name, loc.county, loc.parish, count(loc.parish) AS num_holdings_in_parish
    		    FROM (
    				       SELECT DISTINCT cph, county, parish, holding, county_name FROM locations_extended_tbl
    			       ) loc
    		    GROUP BY loc.county, loc.parish, loc.county_name
    		    ORDER BY loc.county, loc.parish
    	    ) loc2
     GROUP BY loc2.county, loc2.county_name
     ORDER BY loc2.county"
    )

```

```{r Save: Counties structure, eval = FALSE, echo = FALSE}

saveRDS(df_count_parishes_and_holdings_per_county, 'Tables/df_count_parishes_and_holdings_per_county.rds')

```

```{r Read: Counties structure}

df_count_parishes_and_holdings_per_county <- read_rds('Tables/df_count_parishes_and_holdings_per_county.rds')

df_count_parishes_and_holdings_per_county

```




## Premesis Types

### The number of each premesis type in the data:

```{r Query: Premesis Type, eval = FALSE}

df_count_premesis_type = dbGetQuery(con, 
    "SELECT loc.premises_type, count(loc.premises_type) AS num_premises_type
     FROM (
		        SELECT DISTINCT cph, county, parish, holding, county_name, premises_type FROM locations_extended_tbl
	        ) loc
    GROUP BY loc.premises_type
    ORDER BY num_premises_type DESC"
    )

```

```{r Save: Premesis Type, eval = FALSE, echo = FALSE}

saveRDS(df_count_premesis_type, 'Tables/df_count_premesis_type.rds')

```

```{r Read: Premesis Type}

df_count_premesis_type <- read_rds('Tables/df_count_premesis_type.rds')

df_count_premesis_type

```


### The number of each premesis type in Cheshire:

```{r Query: Premesis Type Cheshire, eval = FALSE}

df_count_premesis_type_cheshire = dbGetQuery(con, 
    "SELECT loc.county_name, loc.premises_type, count(loc.premises_type) AS num_premises_type
     FROM (
		        SELECT DISTINCT cph, county, parish, holding, county_name, premises_type FROM locations_extended_tbl
		        WHERE county_name = 'CHESHIRE'
	        ) loc
     GROUP BY loc.premises_type, loc.county_name
     ORDER BY num_premises_type DESC"
    )

```

```{r Save: Premesis Type Cheshire, eval = FALSE, echo = FALSE}

saveRDS(df_count_premesis_type_cheshire, 'Tables/df_count_premesis_type_cheshire.rds')

```

```{r Read: Premesis Type Cheshire}

df_count_premesis_type_cheshire <- read_rds('Tables/df_count_premesis_type_cheshire.rds')

df_count_premesis_type_cheshire

```


# Herds

## The number of unique cattle in the UK all time

```{r Query: Unique Cattle, eval = FALSE}

df_count_unique_cattle = dbGetQuery(con, 
    "SELECT eartag, count(eartag) AS count_eartag
     FROM animal_details_v2_tbl
     GROUP BY eartag
     ORDER BY count_eartag DESC
     "
    )

```

```{r Save: Unique Cattle, eval = FALSE, echo = FALSE}

saveRDS(df_count_unique_cattle, 'Tables/df_count_unique_cattle.rds')

```

```{r Read: Unique Cattle}

df_count_unique_cattle <- read_rds('Tables/df_count_unique_cattle.rds')

df_count_unique_cattle

```


## The number of unique cattle in Cheshire all time

```{r Query: Unique Cattle Cheshire, eval = FALSE}

# TO FIX: Query

df_count_unique_cattle_cheshire = dbGetQuery(con, 
    "SELECT ADv2.eartag, ADv2.dob, ADv2.dod,
            Mext.movement_date, Mext.off_cph, Mext.on_cph, Mext.birth, Mext.death, Mext.off_county_name, Mext.on_county_name 
     FROM animal_details_v2_tbl ADv2
      INNER JOIN movement_extended_tbl Mext ON (Mext.eartag = ADv2.eartag)
     WHERE Mext.off_county_name = 'CHESHIRE' OR Mext.on_county_name = 'CHESHIRE'"
    )

```

```{r Save: Unique Cattle Cheshire, eval = FALSE, echo = FALSE}

saveRDS(df_count_unique_cattle_cheshire, 'Tables/df_count_unique_cattle_cheshire.rds')

```

```{r Read: Unique Cattle Cheshire}

df_count_unique_cattle_cheshire <- read_rds('Tables/df_count_unique_cattle_cheshire.rds')

df_count_unique_cattle_cheshire

```



# Movements

## The number of unique movements in the UK all time

```{r Query: Unique Movements, eval = FALSE}

df_count_unique_movements = dbGetQuery(con, 
    "SELECT COUNT(*) as total_num_movements_all_time
     FROM moves_no_dups_with_sameday_death_tbl
     "
    )

```

```{r Save: Unique Movements, eval = FALSE, echo = FALSE}

saveRDS(df_count_unique_movements, 'Tables/df_count_unique_movements.rds')

```

```{r Read: Unique Movements}

df_count_unique_movements <- read_rds('Tables/df_count_unique_movements.rds')

df_count_unique_movements

```


## The number of unique animals moved in the UK all time

```{r Query: Unique Animals Moved, eval = FALSE}

df_count_unique_animals_moved = dbGetQuery(con, 
    "SELECT COUNT(DISTINCT eartag) as total_num_unique_animals_moved_all_time
     FROM moves_no_dups_with_sameday_death_tbl
     "
    )

```

```{r Save: Unique Animals Moved, eval = FALSE, echo = FALSE}

saveRDS(df_count_unique_animals_moved, 'Tables/df_count_unique_animals_moved.rds')

```

```{r Read: Unique Animals Moved}

df_count_unique_animals_moved <- read_rds('Tables/df_count_unique_animals_moved.rds')

df_count_unique_animals_moved

```




## The number of unique births in the UK all time

```{r Query: Unique Births, eval = FALSE}

df_count_unique_births = dbGetQuery(con, 
    "SELECT COUNT(*) as total_num_births_all_time
     FROM births_no_dups_tbl
     "
    )

```

```{r Save: Unique Births, eval = FALSE, echo = FALSE}

saveRDS(df_count_unique_births, 'Tables/df_count_unique_births.rds')

```

```{r Read: Unique Births}

df_count_unique_births <- read_rds('Tables/df_count_unique_births.rds')

df_count_unique_births

```



## The number of unique deaths in the UK all time

```{r Query: Unique Deaths, eval = FALSE}

df_count_unique_deaths = dbGetQuery(con, 
    "SELECT COUNT(*) as total_num_deaths_all_time
     FROM deaths_no_dups_tbl
     "
    )

```

```{r Save: Unique Deaths, eval = FALSE, echo = FALSE}

saveRDS(df_count_unique_deaths, 'Tables/df_count_unique_deaths.rds')

```

```{r Read: Unique Deaths}

df_count_unique_deaths <- read_rds('Tables/df_count_unique_deaths.rds')

df_count_unique_deaths

```

## The unique movements into or out of Cheshire all time

```{r Query: Unique Movements Cheshire, eval = FALSE}

df_cheshire_unique_movements = dbGetQuery(con, 
    "SELECT *
     FROM moves_no_dups_with_sameday_death_tbl
     WHERE off_county_name = 'CHESHIRE' OR on_county_name = 'CHESHIRE'
     "
    )

```

```{r Save: Unique Movements Cheshire, eval = FALSE, echo = FALSE}

saveRDS(df_cheshire_unique_movements, 'Tables/df_cheshire_unique_movements.rds')

```

```{r Read: Unique Movements Cheshire}

df_cheshire_unique_movements <- read_rds('Tables/df_cheshire_unique_movements.rds')

df_cheshire_unique_movements

```


## The unique births in Cheshire all time

```{r Query: Unique Births Cheshire, eval = FALSE}

df_cheshire_unique_births = dbGetQuery(con, 
    "SELECT *
     FROM births_no_dups_tbl
     WHERE off_county_name = 'CHESHIRE' OR on_county_name = 'CHESHIRE'
     "
    )

```

```{r Save: Unique Births Cheshire, eval = FALSE, echo = FALSE}

saveRDS(df_cheshire_unique_births, 'Tables/df_cheshire_unique_births.rds')

```

```{r Read: Unique Births Cheshire}

df_cheshire_unique_births <- read_rds('Tables/df_cheshire_unique_births.rds')

df_cheshire_unique_births

```


## The unique deaths in Cheshire all time

```{r Query: Unique Deaths Cheshire, eval = FALSE}

df_cheshire_unique_deaths = dbGetQuery(con, 
    "SELECT *
     FROM deaths_no_dups_tbl
     WHERE off_county_name = 'CHESHIRE' OR on_county_name = 'CHESHIRE'
     "
    )

```

```{r Save: Unique Deaths Cheshire, eval = FALSE, echo = FALSE}

saveRDS(df_cheshire_unique_deaths, 'Tables/df_cheshire_unique_deaths.rds')

```

```{r Read: Unique Deaths Cheshire}

df_cheshire_unique_deaths <- read_rds('Tables/df_cheshire_unique_deaths.rds')

df_cheshire_unique_deaths

```


## The number of unique animals moved into or out of Cheshire all time

```{r Query: Unique Animals Moved Cheshire, eval = FALSE}

df_cheshire_count_unique_animals_moved = dbGetQuery(con, 
    "SELECT COUNT(DISTINCT eartag) as total_num_unique_animals_moved_all_time
     FROM moves_no_dups_with_sameday_death_tbl
     WHERE off_county_name = 'CHESHIRE' OR on_county_name = 'CHESHIRE'
     "
    )

```

```{r Save: Unique Animals Moved Cheshire, eval = FALSE, echo = FALSE}

saveRDS(df_cheshire_count_unique_animals_moved, 'Tables/df_cheshire_count_unique_animals_moved.rds')

```

```{r Read: Unique Animals Moved Cheshire}

df_cheshire_count_unique_animals_moved <- read_rds('Tables/df_cheshire_count_unique_animals_moved.rds')

df_cheshire_count_unique_animals_moved

```





# Tests

## The number of unique tests in the UK all time

```{r Query: Unique Tests, eval = FALSE}

df_count_unique_tests = dbGetQuery(con, 
    "SELECT COUNT(*) as total_num_tests_all_time
     FROM tests_no_dups
     "
    )

```

```{r Save: Unique Tests, eval = FALSE, echo = FALSE}

saveRDS(df_count_unique_tests, 'Tables/df_count_unique_tests.rds')

```

```{r Read: Unique Tests}

df_count_unique_tests <- read_rds('Tables/df_count_unique_tests.rds')

df_count_unique_tests

```

## The number of unique tests in each category in the UK all time

```{r Query: Unique Test Category, eval = FALSE}

df_count_unique_test_category = dbGetQuery(con, 
    "SELECT category, COUNT(category) as total_num_tests_all_time
     FROM tests_no_dups
     GROUP BY category
     ORDER BY total_num_tests_all_time DESC
     "
    )

```

```{r Save: Unique Tests Category, eval = FALSE, echo = FALSE}

saveRDS(df_count_unique_test_category, 'Tables/df_count_unique_test_category.rds')

```

```{r Read: Unique Tests Category}

df_count_unique_test_category <- read_rds('Tables/df_count_unique_test_category.rds')

df_count_unique_test_category

```

## The number of unique tests for each reason in the UK all time

```{r Query: Unique Test Reason, eval = FALSE}

df_count_unique_test_type = dbGetQuery(con, 
    "SELECT test_type, COUNT(test_type) as total_num_tests_all_time
     FROM tests_no_dups
     GROUP BY test_type
     ORDER BY total_num_tests_all_time DESC
     "
    )

```

```{r Save: Unique Tests Reason, eval = FALSE, echo = FALSE}

saveRDS(df_count_unique_test_type, 'Tables/df_count_unique_test_type.rds')

```

```{r Read: Unique Tests Reason}

df_count_unique_test_type <- read_rds('Tables/df_count_unique_test_type.rds')

df_count_unique_test_type

```


## The number of unique animals tested in the UK all time

```{r Query: Unique Animals Tested, eval = FALSE}

df_count_unique_animals_tested = dbGetQuery(con, 
    "SELECT COUNT(DISTINCT eartag) as total_num_unique_animals_tested_all_time
     FROM tests_no_dups
     "
    )

```

```{r Save: Unique Animals Tested, eval = FALSE, echo = FALSE}

saveRDS(df_count_unique_animals_tested, 'Tables/df_count_unique_animals_tested.rds')

```

```{r Read: Unique Animals Tested}

df_count_unique_animals_tested <- read_rds('Tables/df_count_unique_animals_tested.rds')

df_count_unique_animals_tested

```


## The tests in Cheshire all time

```{r Query: Cheshire tests, eval = FALSE}

df_cheshire_tests = dbGetQuery(con, 
    "SELECT *
     FROM tests_no_dups
     WHERE county_name = 'CHESHIRE'
     "
    )

```

```{r Save: Cheshire tests, eval = FALSE, echo = FALSE}

saveRDS(df_cheshire_tests, 'Tables/df_cheshire_tests.rds')

```

```{r Read: Cheshire tests}

df_cheshire_tests <- read_rds('Tables/df_cheshire_tests.rds')

df_cheshire_tests

```


## The number of unique tests in each category in Chesire all time

```{r Query: Unique Test Category Cheshire, eval = FALSE}

df_count_unique_test_category_cheshire = dbGetQuery(con, 
    "SELECT category, COUNT(category) as total_num_tests_all_time
     FROM (SELECT *
           FROM tests_no_dups
           WHERE county_name = 'CHESHIRE') as top
     GROUP BY category
     ORDER BY total_num_tests_all_time DESC
     "
    )

```

```{r Save: Unique Tests Category Cheshire, eval = FALSE, echo = FALSE}

saveRDS(df_count_unique_test_category_cheshire, 'Tables/df_count_unique_test_category_cheshire.rds')

```

```{r Read: Unique Tests Category Cheshire}

df_count_unique_test_category_cheshire <- read_rds('Tables/df_count_unique_test_category_cheshire.rds')

df_count_unique_test_category_cheshire

```

## The number of unique tests for each reason in Cheshire all time

```{r Query: Unique Test Reason Cheshire, eval = FALSE}

df_count_unique_test_type_cheshire = dbGetQuery(con, 
    "SELECT test_type, COUNT(test_type) as total_num_tests_all_time
     FROM (SELECT *
           FROM tests_no_dups
           WHERE county_name = 'CHESHIRE') as top
     GROUP BY test_type
     ORDER BY total_num_tests_all_time DESC
     "
    )

```

```{r Save: Unique Tests Reason Cheshire, eval = FALSE, echo = FALSE}

saveRDS(df_count_unique_test_type_cheshire, 'Tables/df_count_unique_test_type_cheshire.rds')

```

```{r Read: Unique Tests Reason Cheshire}

df_count_unique_test_type_cheshire <- read_rds('Tables/df_count_unique_test_type_cheshire.rds')

df_count_unique_test_type_cheshire

```

## The number of unique animals tested in Cheshire all time

```{r Query: Unique Animals Tested Cheshire, eval = FALSE}

df_count_unique_animals_tested_cheshire = dbGetQuery(con, 
    "SELECT COUNT(DISTINCT eartag) as total_num_unique_animals_tested_all_time
     FROM (SELECT *
           FROM tests_no_dups
           WHERE county_name = 'CHESHIRE') as top
     "
    )

```

```{r Save: Unique Animals Tested Cheshire, eval = FALSE, echo = FALSE}

saveRDS(df_count_unique_animals_tested_cheshire, 'Tables/df_count_unique_animals_tested_cheshire.rds')

```

```{r Read: Unique Animals Tested Cheshire}

df_count_unique_animals_tested_cheshire <- read_rds('Tables/df_count_unique_animals_tested_cheshire.rds')

df_count_unique_animals_tested_cheshire

```










